---
title: "Stormwater Denitrification"
author: "Ariane L. Peralta, Mario E. Muscarella, Eban Bean"
date: "Last updated on `r format(Sys.time(), '%d %B, %Y')`"
header-includes:
  - \usepackage{array}
  - \usepackage{graphics}
  - \usepackage[utf8]{inputenc}
output:
  pdf_document:
  fig_caption: true
---

Project Description:
Is there a phylogenetic diversity-functioning relationship in this urban stream ecosystem? 
Does magnitude of function influence diversity-function relationship?
Does sample type (sed vs. water) matter?

ToDo
1. Calculate relative abundance for each tip
2. Regress relative abundance against denitrification
3. Generate all sub-trees
4. Loop through sub-trees and add relative abundances for tips in sub tree
5. Regress subtree relative abundance against denitrification


# Initial Setup
```{r, results="hide", message=FALSE}
rm(list=ls())
setwd("~/GitHub/StormwaterDenitrification/analyses")
se <- function(x, ...){sd(x, na.rm = TRUE)/sqrt(length(na.omit(x)))}
ci <- function(x, ...){1.96 * sd(x,na.rm = TRUE)}

# Code Dependencies
source("../bin/MothurTools.R")
require("vegan")
require("reshape2")
require("dplyr")
require("ggplot2")
require("nlme")
require("ade4")
require("grid"); require("png")
require("ape"); require("picante")
library("agricolae")

myColors <- c("#448844", "#33CC33", "#CCFF00", "#FFF000", "#FF9933", "#A9A9A9")
#names(myColors) <- c()

# Calculates Simpsons Evenness
simp_even <- function(SAD = " "){
  SAD <- subset(SAD, SAD > 0)
  S <- length(SAD)
  N <- sum(SAD)
  X <- rep(NA, S)
  for (i in 1:S){
    X[i] <- (SAD[i]*(SAD[i] - 1)) / (N * (N - 1))
    }
  D <- sum(X)
  e_d <- (1/D)/S
  return(e_d)
  }
```

# Import Data Files
## Experimental Design File
```{r}
# Import Environmental Data
design <- read.delim("../data/TC.design.txt", row.names = 1)
```

## Microbial Data
```{r}
# Import OTU data
# Import Raw Data
otu.in <- read.otu("../data/TC.bac.final.shared")

# Correct Sample IDs and Subset Original Design File
rownames(design) <- gsub("_0", "_", gsub("^ECU_", "", rownames(design)))
design.otu <- design[which(rownames(design) %in% rownames(otu.in)), ]
all.equal(rownames(otu.in), rownames(design.otu))

# Remove OTUs with less than two occurences across all sites
otus <- otu.in[, which(colSums(otu.in) >= 2)]

# Make Presence Absence Matrix
dataPA <- (otus > 0) * 1

# Make Relative Abundence Matrices
dataREL <- otus
for(i in 1:dim(otus)[1]){
  dataREL[i,] <- otus[i,]/sum(otus[i,])
}

# Log Transform Relative Abundances
dataREL.log <- decostand(dataREL, method="log")

# Import Taxonomy File
otu.tax <- read.tax(taxonomy = "../data/TC.bac.final.0.03.taxonomy",
                   format = "rdp", tax.levels = 6, col.tax = 3)
```

# Simple Hypothesis Testing Bacterial Community Composition - 
```{r}
#PERMANOVA on baseflow only
new.data <-cbind(design.otu,dataREL)
as.data.frame(new.data)
new.data.2 <- subset(new.data, Storm == "baseflow")

adonis = adonis(new.data.2[,-c(1:6)] ~ Sample_Type, method = "bray", data = new.data.2, perm=1000)
adonis
```
# Microbial Ordinations

## Principal Coordinates Ordination
```{r}
# TEST with original labels
# Create Distance Matrix
sampleREL.dist <- vegdist(dataREL, method="bray")

# Principal Coordinates Analysis
TC_pcoa <- cmdscale(sampleREL.dist, k=3, eig=TRUE, add=FALSE)
  # Classical (Metric) Multidimensional Scaling; returns PCoA coordinates
  # eig=TRUE returns eigenvalues; k = # of dimensions to calculate

explainvar1 <- round(TC_pcoa$eig[1] / sum(TC_pcoa$eig), 3) * 100
explainvar2 <- round(TC_pcoa$eig[2] / sum(TC_pcoa$eig), 3) * 100
sum.eig <- sum(explainvar1, explainvar2)

points <- cbind(as.data.frame(TC_pcoa$points), design.otu$Sample_Type)
L.centroids <- melt(points, id="design.otu$Sample_Type", measure.vars = c("V1", "V2"))
centroids <- cast(L.centroids, variable ~ design.otu$Sample_Type, mean)
centroids.se <- cast(L.centroids, variable ~ design.otu$Sample_Type, se)
centroids.sd <- cast(L.centroids, variable ~ design.otu$Sample_Type, sd)

cent.dataframe <- t(data.frame(rbind(centroids[1,-1], centroids[2,-1],
                             centroids.sd[1,-1],centroids.sd[2,-1])))
colnames(cent.dataframe) <- c("V1", "V2", "V1e", "V2e")
cent.treats <- rownames(cent.dataframe)

explainvar1 #Include in PCoA 1 axis label 18.7%
explainvar2 #Include in PCoA 2 axis label 16.5%
                                        
df <- as.data.frame(cent.dataframe)
p <- ggplot(df, aes(x=V1, y=V2, colour=cent.treats)) + theme_bw() 
p + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + 
theme(panel.background = element_blank()) + 
  geom_errorbarh(aes(xmax=V1+V1e, xmin=V1-V1e, height=0.01), colour="black") +    
  geom_errorbar(aes(ymax=V2+V2e, ymin=V2-V2e, width=0.01), colour="black") + 
  geom_point(size=5) 

```

```{r}
# Create Distance Matrix
sampleREL.dist <- vegdist(BDdataREL, method="bray")

# Principal Coordinates Analysis
TC_pcoa <- cmdscale(sampleREL.dist, k=3, eig=TRUE, add=FALSE)
  # Classical (Metric) Multidimensional Scaling; returns PCoA coordinates
  # eig=TRUE returns eigenvalues; k = # of dimensions to calculate

explainvar1 <- round(TC_pcoa$eig[1] / sum(TC_pcoa$eig), 3) * 100
explainvar2 <- round(TC_pcoa$eig[2] / sum(TC_pcoa$eig), 3) * 100
sum.eig <- sum(explainvar1, explainvar2)
```

# Plot
```{r}
treatments <- factor(design.otu$Sample_Type, levels = c("BD", "BK", 
                     "W"))
points <- cbind(as.data.frame(TC_pcoa$points), treatments)
L.centroids <- melt(points, id="treatments", measure.vars = c("V1", "V2"))
centroids <- cast(L.centroids, variable ~ treatments, mean)
centroids.se <- cast(L.centroids, variable ~ treatments, se)
centroids.sd <- cast(L.centroids, variable ~ treatments, sd)

cent.dataframe <- t(data.frame(rbind(centroids[1,-1], centroids[2,-1],
                             centroids.sd[1,-1],centroids.sd[2,-1])))
colnames(cent.dataframe) <- c("V1", "V2", "V1e", "V2e")
cent.treats <- rownames(cent.dataframe)

explainvar1 #Include in PCoA 1 axis label 18.7%
explainvar2 #Include in PCoA 2 axis label 16.5%

###need to incorporate shape for Location and color for Sample_Type

pcoa.groups <- paste(design.otu$Sample_Type, design.otu$Location, sep = "_")
pcoa.points <- data.frame(TC_pcoa$points, group = pcoa.groups)

# Calculate Centroids (mean and SE)
pcoa.L.centroids <- melt(pcoa.points, id="group", measure.vars = c("X1", "X2"))
pcoa.centroids <- acast(pcoa.L.centroids, variable ~ group, mean)
pcoa.centroids.se <- acast(pcoa.L.centroids, variable ~ group, se)
pcoa.centroids.sd <- acast(pcoa.L.centroids, variable ~ group, sd)

# Combine
pcoa.cent.dataframe <- cbind(t(pcoa.centroids), t(pcoa.centroids.se))
colnames(pcoa.cent.dataframe) <- c("V1", "V2", "V1e", "V2e")
pcoa.cent.treats <- rownames(pcoa.cent.dataframe)


#STUCK HERE
pcoa.col <- as.factor(sapply(strsplit(pcoa.cent.treats, "_"), `[`, 2)) # Salinity
pcoa.shape <- as.factor(sapply(strsplit(pcoa.cent.treats, "_"), `[`, 1))  # Date2



df <- as.data.frame(cent.dataframe)
p <- ggplot(df, aes(x=V1, y=V2, colour=cent.treats)) + theme_bw() 
p + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + 
theme(panel.background = element_blank()) + 
  geom_errorbarh(aes(xmax=V1+V1e, xmin=V1-V1e, height=0.01), colour="black") +    
  geom_errorbar(aes(ymax=V2+V2e, ymin=V2-V2e, width=0.01), colour="black") + 
  geom_point(size=5,shape=as.factor(design.otu$Location)) +
  scale_colour_manual(labels = c("stream bed", "stream bank", "water"), values = c("brown","orange","blue")) +
  scale_shape_manual(labels = c("B","C","D"), values = c(22, 21, 24)) +
theme(axis.title=element_text(size=18), axis.text.x=element_text(size=14), panel.border = element_rect(colour = "black",size=1.25)) + 
  theme(axis.ticks.length=unit(0.3,"cm")) + 
  xlab("PCoA 1 (18.7%)") + ylab("PCoA 2 (16.5%)") + 
  labs(fill = "Sample Type", shape = "Location") +
  guides(fill = guide_legend(override.aes = list(pch=21, size = 4, colour="black")),
         shape = guide_legend(override.aes = list(size = 4, fill="black")))
  
ggsave("../figures/16SrRNA_BCC_Rplot.pdf", plot=last_plot(), device=NULL, path=NULL, scale=1, width=NA, height=NA, dpi=300, limitsize=TRUE)
```

## Phylogenetic Tree and UniFrac Distance Matrix
```{r}
# Import Tree
tree <- read.tree("../data/TC.bac.rename.tree")

tree$tip.label[1:5]
tree$tip.label <- gsub("\\|", "", tree$tip.label)

# Import Unifrac Matrix
unifrac.raw <- read.delim("../data/TC.bac.tree1.weighted.phylip.dist",
                             header = F, skip = 1, row.names = 1)

row.names(unifrac.raw) <- gsub("     ", "", row.names(unifrac.raw))

unifrac <- unifrac.raw[which(row.names(unifrac.raw) %in%
                                   row.names(otus)),
                             which(row.names(unifrac.raw) %in%
                                   row.names(otus))]

rownames(unifrac) <- gsub("     ", "", row.names(unifrac))
colnames(unifrac) <- rownames(unifrac)
dim(unifrac)

# Make into Distance Matrix
unifrac.dist <- as.dist(unifrac, upper = T, diag = T)
```

## Notes: Phylogenetic Analysis
```{sh, eval = F}
# The following was done outside of R
python ../bin/name_change.py TC.bac.final.0.03.rep.fasta TC.bac.final.0.03.rep.rename.fasta

# This version is needed for Mothur
FastTree -gtr -nt -gamma -fastest TC.bac.final.0.03.rep.fasta > TC.bac.tree

# This version is needed for R analysis
FastTree -gtr -nt -gamma -fastest TC.bac.final.0.03.rep.rename.fasta > TC.bac.rename.tree

Mothur (v 1.39.5)
system(cp ./TC.trim.contigs.good.unique.good.filter.good.unique.precluster.pick.pick.opti_mcc.unique_list.0.03.rep.fasta ./TC.bac.final.0.03.rep.fasta)
system(cp ./TC.trim.contigs.good.unique.good.filter.good.unique.precluster.pick.pick.opti_mcc.unique_list.0.03.rep.count_table ./TC.bac.final.count)
unifrac.weighted(tree=TC.bac.tree, count=TC.bac.final.count, distance=square, processors=8)

Output File Names:
TC.bac.treewsummary
TC.bac.tree1.weighted.phylip.dist
```

## Denitrification
```{r}
# Import Data
dat <- read.delim("../data/WQ_DEA_separate/2015_TC_DEArate.txt")
dat$Replicate <- as.factor(dat$Replicate)
```

# Sediment Denitrification
```{r}
dat.sed <- dat[dat$Type != "W", ]

dim1 <- length(dat.sed$acetyleneb[dat.sed$acetyleneb == "-"])

sed.eff <- as.data.frame(matrix(NA, dim1, 4))
colnames(sed.eff) <- c("Location", "Type", "Replicate", "Rate")
sed.eff$Location <- dat.sed$Location[dat.sed$acetyleneb == "-"]
sed.eff$Type <- dat.sed$Type[dat.sed$acetyleneb == "-"]
sed.eff$Replicate <- dat.sed$Replicate[dat.sed$acetyleneb == "-"]
sed.eff$Rate <- (dat.sed$Rate[dat.sed$acetyleneb == "+"] -
					   dat.sed$Rate[dat.sed$acetyleneb == "-"] )

# Reorganize
sed.eff.m <- melt(sed.eff)
sed.eff.c <- cast(data = sed.eff.m, Location + Type ~ variable, c(mean, se), na.rm=T)

sed.eff.c <- as.data.frame(sed.eff.c)
sed.eff.m <- as.data.frame(sed.eff.m)

# # ANOVA - update
aov.sed=aov(value~Location*Type,data=sed.eff.m)
summary(aov.sed)
posthocD <-TukeyHSD(aov.sed,"Location:Type",conf.level=0.95)
posthocD
```

# Sediment Denitrification Plot
```{r}
# Plot - Sediment N2O
png(filename="../figures/SedN2Oonly.png",
    width = 1200, height = 800, res = 96*2)

pdf(file = "../figures/SedN2Oonly.pdf", width=6, height=4, paper='special')

par(mar=c(3,6,0.5,0.5), oma=c(1,1,1,1)+0.1, lwd=2)
bp_plot <- barplot(sed.eff.c[,3],
					ylab = "", xlab = "",
					lwd=3, col="gray", cex.lab=1.25, cex.names = 1.25,
					ylim = c(-2, 60), yaxt="n",
          space = c(1, 0.25, 1, 0.25, 1, 0.25),
          density=c(-1, 15, -1, 15, -1, 15))
arrows(x0 = bp_plot, y0 = sed.eff.c[,3], y1 = sed.eff.c[,3] - sed.eff.c[,4],
       angle = 90, length=0.1, lwd = 2)
arrows(x0 = bp_plot, y0 = sed.eff.c[,3], y1 = sed.eff.c[,3] + sed.eff.c[,4],
       angle = 90, length=0.1, lwd = 2)
axis(side = 2, labels=T, lwd.ticks=2, las=2, lwd=2, at = c(seq(0, 60, 20)))
mtext(c("Downstream\nSeep", "Stream\nMiddle", "Culvert\n"),
      side = 1, at=c(2.2, 5.4, 8.6),
      line = 1.5, cex=1, adj=0.5)
mtext("Denitrification Rate", side = 2, line = 3.25, cex = 1.25)
mtext(side = 2, line = 2,  text = expression(paste("(ng N"[2], "O g"^-1, "sediment hr"^-1, ")")))
abline(h=0, lwd=2, lty=3)
legend("topright", c("stream bed", "stream bank"), fill="gray", bty="n", cex=1,
       density=c(-1, 15))

dev.off() # sthis writes plot to folder
graphics.off()

## Show Plot
img <- readPNG("../figures/SedN2Oonly.png")
grid.raster(img)
```

# Water Denitrification
```{r}
dat.water <- dat[dat$Type == "W", ]

# Deal with negative values
#dat.water[,7:10][dat.water[,7:10] < 0] <- 0

dim1 <- length(dat.water$acetyleneb[dat.water$acetyleneb == "-"])

wtr.eff <- as.data.frame(matrix(NA, dim1, 4))
colnames(wtr.eff) <- c("Location", "Time", "Replicate", "Rate")
wtr.eff$Location <- dat.water$Location[dat.water$acetyleneb == "-"]
wtr.eff$Time <- dat.water$Time[dat.water$acetyleneb == "-"]
wtr.eff$Replicate <- dat.water$Replicate[dat.water$acetyleneb == "-"]
wtr.eff$Rate <- (dat.water$Rate[dat.water$acetyleneb == "+"] -
					   dat.water$Rate[dat.water$acetyleneb == "-"] )

# Reorganize Data
wtr.eff.m <- melt(wtr.eff)
wtr.eff.c <- cast(data = wtr.eff.m, Location + Time ~ variable, c(mean, se), na.rm=T)

wtr.eff.c <- as.data.frame(wtr.eff.c)
wtr.eff.m <- as.data.frame(wtr.eff.m)

# # ANOVA - update
aov.wtr=aov(value~Location*Time,data=wtr.eff.m)
summary(aov.wtr)

# Df Sum Sq Mean Sq F value Pr(>F)
# Location       5 0.7579 0.15157   1.544  0.221
# Time           1 0.1136 0.11360   1.157  0.295
# Location:Time  3 0.5062 0.16872   1.718  0.195
# Residuals     20 1.9638 0.09819
```

# Water Denitrification Plot
```{r}
png(filename="../figures/WaterN2Oonly.png",
    width = 1200, height = 800, res = 96*2)

pdf(file = "../figures/WaterN2Oonly.pdf", width=6, height=4, paper='special')

par(mar=c(3,6,0.5,0.5), oma=c(1,1,1,1)+0.1, lwd=2)
bp_plot <- barplot(wtr.eff.c[,3],
					ylab = "",
					lwd=3, yaxt="n", col="gray", ex.lab=1.5, cex.names = 1.25,
					ylim = c(-1, 2),
                   	space = c(1, 0.25, 1, 0.25, 1, 0.25, 1, 0.25, 1, 1),
                   	density=c(-1, 15, -1, 15, -1, 15, -1, 15, 15, 15))
arrows(x0 = bp_plot, y0 = wtr.eff.c[,3], y1 = wtr.eff.c[,3] - wtr.eff.c[,4], angle = 90,
       length=0.1, lwd = 2)
arrows(x0 = bp_plot, y0 = wtr.eff.c[,3], y1 = wtr.eff.c[,3] + wtr.eff.c[,4], angle = 90,
       length=0.1, lwd = 2)
axis(side = 2, labels=T, lwd.ticks=2, las=2, lwd=2, at = c(seq(-1, 2, 1)))
mtext(c("Tar River\nOutflow", "Downstream\nSeep", "Stream\nMiddle", "Culvert\n",
        "Upstream\nCulvert", "Upstream\nInflow"),
      side = 1, at=c(2, 5.5, 9, 11.75, 14.5, 17),
      line = 2, cex=0.8, adj=0.5)
mtext("Denitrification Rate", side = 2, line = 3.25, cex = 1.25)
mtext(side = 2, line = 2,  text = expression(paste("(ng N"[2], "O g"^-1, "sediment hr"^-1, ")")))

abline(h=0, lwd=2, lty=3)
legend("topright", c("baseflow", "stormflow"), fill="gray", bty="n", cex=1,
       density=c(-1, 15))


dev.off() # this writes plot to folder
graphics.off() # shuts down open devices

## Show Plot
img <- readPNG("../figures/WaterN2Oonly.png")
grid.raster(img)
```

# Diversity-Function Relationships
```{r}
min(rowSums(otus))

# remove low count sites
bad <- which(rowSums(otus) < 10000)
otus.2 <- otus[-c(bad), ]
otus.2 <- otus.2[, -c(which(colSums(otus.2) == 0))]
min(rowSums(otus.2))

otus.r <- rrarefy(otus.2, 10000)

# Fisher's Alpha
fisher <- fisher.alpha(otus.r)
fisher

# Species Richness
richness <- rowSums((otus.r >= 1))

# Shannon Diversity (my function gets the same answer)
shannon <- diversity(otus.r, "shannon")

# Simpson's Evenness
simp.even <- apply(otus.r, 1, simp_even)

# Phylogenetic Diversity
# Reorder Tree
which(tree$edge.length < 0.0000001)
tree.2 <- reorder(tree, order = "cladewise")
tree.r <- midpoint.root(tree.2)
phylo.d <- pd(otus.r, tree, include.root = F)

design.r <- design.otu[which(rownames(design.otu) %in% rownames(otus.r)), ]


design.sed <- design.r[design.r$Sample_Type != "W", ]

sed.bac <- as.data.frame(matrix(NA, dim(design.sed)[1], 9))
colnames(sed.bac) <- c("Sample", "Location", "Time", "Type", "Replicate", "S", "F_alpha", 
                       "S_even", "Shannon")
sed.bac$Sample <- rownames(design.sed)
sed.bac$Location <- design.sed$Location
sed.bac$Time <- design.sed$Storm
sed.bac$Type <- design.sed$Sample_Type
sed.bac$Replicate <- design.sed$Field_replicate
for(i in 1:length(rownames(design.sed))){
  temp <- rownames(design.sed)[i]
  sed.bac$S[i] <- richness[which(names(richness) == temp)]
  sed.bac$F_alpha[i] <- fisher[which(names(fisher) == temp)]
  sed.bac$S_even[i] <- simp.even[which(names(simp.even) == temp)]
  sed.bac$Shannon[i] <- shannon[which(names(shannon) == temp)]
}

# Reorganize
sed.bac.m <- melt(sed.bac, id = c("Sample", "Location", "Time", "Type", "Replicate"))
sed.bac.c <- cast(data = sed.bac.m, Location + Type + Replicate ~ variable, c(mean, se), na.rm=T)

sed.bac.c <- as.data.frame(sed.bac.c)
sed.bac.m <- as.data.frame(sed.bac.m)

# # ANOVA - update
aov.sed=aov(value~Location*Type,data=sed.bac.m[which(sed.bac.m$variable == "S"), ])
summary(aov.sed)
posthocD <-TukeyHSD(aov.sed,"Location:Type",conf.level=0.95)
posthocD

```








# Water Quality Data
```{r}
layout(matrix(c(1,2, 3), 1, 3, byrow = TRUE), widths = c(2, 2, 2))
## show the regions that have been allocated to each plot
layout.show(3)

# Import Data - water quality
chem <- read.delim("../data/WQ_DEA_separate/2015_TC_WaterQualityData.txt")
chem$Replicate <- as.factor(chem$Replicate)

chem.m <- melt(chem, id.vars = c("Location", "Type", "Storm"), measure.vars = "NH4")
chem.c <- cast(data = chem.m, Location + Storm  ~ variable, c(mean, se), na.rm=T)

chem.c <- as.data.frame(chem.c)
```

# Water quality Plot
```{r}
png(filename="../figures/WaterQuality.png",
    width = 1200, height = 800, res = 96*2)

par(mar=c(3,6,0.5,0.5), oma=c(1,1,1,1)+0.1, lwd=2)
TCplot <- barplot(chem.c[,3], ylab = "Ammonium (mg/L)",
                   lwd=3, yaxt="n", col="gray",
                   cex.lab=1.5, cex.names = 1.25, 
                  space = c(1, 0.25, 1, 0.25, 1, 0.25, 1, 0.25, 1, 1, 1, 1, 1))

points(TCplot, chem.c[,3], pch=22, cex = 2, density=c(-1, 15, -1, 15, -1, 15, -1, 15, 15, 15, 15, 15, 15))
arrows(x0 = TCplot, y0 = chem.c[,3], y1 = chem.c[,3] - chem.c[,4], angle = 90,
       length=0.1, lwd = 2)
arrows(x0 = TCplot, y0 = chem.c[,3], y1 = chem.c[,3] + chem.c[,4], angle = 90,
       length=0.1, lwd = 2)
axis(side = 2, labels=T, lwd.ticks=2, las=2, lwd=2)
mtext(c("A", "B", "C", "D","E", "F", "G", "H", "I"))
      side = 1, at=c(2, 4, 5, 8.5, 12, 14.5, 16.5),
      line = 2, cex=0.8, adj=0.5)
legend("topright", c("Baseline", "Storm"), fill="gray", bty="n", cex=1.25,
       density=c(-1, 15))

dev.off() # this writes plot to folder
graphics.off() # shuts down open devices

## Show Plot
img <- readPNG("../figures/WaterQuality.png")
grid.raster(img)
```

