---
title: "Stormwater Denitrification"
author: "Ariane L. Peralta, Mario E. Muscarella, Eban Bean"
date: "Last updated on `r format(Sys.time(), '%d %B, %Y')`"
header-includes:
  - \usepackage{array}
  - \usepackage{graphics}
  - \usepackage[utf8]{inputenc}
output:
  pdf_document:
  fig_caption: true
---

Project Description:
Is there a phylogenetic diversity-functioning relationship in this urban stream ecosystem? 
Does magnitude of function influence diversity-function relationship?
Does sample type (sed vs. water) matter?

# Initial Setup
```{r, results="hide", message=FALSE}
rm(list=ls())
setwd("~/GitHub/StormwaterDenitrification/analyses")
se <- function(x, ...){sd(x, na.rm = TRUE)/sqrt(length(na.omit(x)))}
ci <- function(x, ...){1.96 * sd(x,na.rm = TRUE)}

# Code Dependencies
source("../bin/MothurTools.R"); source("../bin/DiversityFunctions.R")
require("labdsv"); require("vegan"); require("reshape2")
require("dplyr"); require("ggplot2"); require("nlme")
require("ade4"); require("grid"); require("png")
require("ape"); require("picante")
```
# Water Quality Data
```{r}
layout(matrix(c(1,2, 3), 1, 3, byrow = TRUE), widths = c(2, 2, 2))
## show the regions that have been allocated to each plot
layout.show(3)

# Import Data - water quality
chem <- read.delim("../data/WQ_DEA_separate/2015_TC_WaterQualityData.txt")
chem$Replicate <- as.factor(chem$Replicate)

#plot NH4
p <- ggplot(chem, aes(x=Location, y=NH4, color=as.factor(Storm))) + geom_boxplot() +
      geom_point(aes(color=factor(Storm)), size=2, position = position_jitterdodge()) +   
      scale_color_manual(name="Time", 
                         values=c("cyan", "darkblue"), 
                         labels = c("baseflow", "stormflow")) 
ammonium <- p + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "", 
          y = (expression(paste("(mg/L)")))) + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1)) +ggtitle("Ammonium-N")
ammonium
#ggsave("../figures/water_ammonium_boxplot.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)

#plot NO3.NO2
p <- ggplot(chem, aes(x=Location, y=NO3.NO2, color=as.factor(Storm))) + geom_boxplot() +
      geom_point(aes(color=factor(Storm)), size=2, position = position_jitterdodge()) +   
      scale_color_manual(name="Time", 
                         values=c("cyan", "darkblue"), 
                         labels = c("baseflow", "stormflow")) 
nitrate <- p + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "", 
          y = (expression(paste("")))) + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1))+ggtitle("Ntrate+Nitrite-N")
nitrate
#ggsave("../figures/water_nitrate_boxplot.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)

#plot TDN
p <- ggplot(chem, aes(x=Location, y=TDN, color=as.factor(Storm))) + geom_boxplot() +
      geom_point(aes(color=factor(Storm)), size=2, position = position_jitterdodge()) +   
      scale_color_manual(name="Time", 
                         values=c("cyan", "darkblue"), 
                         labels = c("baseflow", "stormflow")) 
TDN <- p + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "Location", 
          y = (expression(paste("(mg/L)")))) + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1)) +ggtitle("Total Dissolved Nitrogen")
TDN
#ggsave("../figures/water_TDN_boxplot.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)

#plot DOC
p <- ggplot(chem, aes(x=Location, y=DOC, color=as.factor(Storm))) + geom_boxplot() +
      geom_point(aes(color=factor(Storm)), size=2, position = position_jitterdodge()) +   
      scale_color_manual(name="Time", 
                         values=c("cyan", "darkblue"), 
                         labels = c("baseflow", "stormflow")) 
DOC <- p + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "Location", 
          y = (expression(paste("")))) + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1)) +ggtitle("Dissolved Organic Carbon")
DOC
#ggsave("../figures/water_DOC_boxplot.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```
#combine WQ graphs
```{r Loading Files, include=FALSE}
library("cowplot") #combining plots
library("ggpubr")

# remove legends and axes
n1 <-ammonium + theme(legend.position="none") + rremove("x.text")
n2 <- nitrate + theme(legend.position="none") + rremove("x.text")
n3 <- TDN + theme(legend.position="none")
n4 <- DOC + theme(legend.position="none")

# align all plots vertically
plots <- align_plots(n1,n2,n3,n4, align = 'v', axis = 'l')

# combine graphs
n5 <- plot_grid(plots[[1]],plots[[2]],plots[[3]],plots[[4]], labels = c("A","B","C","D"), label_size = 16, ncol=2, align='v') 

legend <- get_legend(ammonium)

ggdraw(plot_grid(plot_grid(n5),
                 plot_grid(NULL, legend, ncol=1),
                 rel_widths=c(1, 0.2)))

ggsave("../figures/WQ_combined.png", plot=last_plot(), device=NULL, path=NULL, scale=1,  width=10, height=6, dpi=300, limitsize=TRUE)
```

# Import Data Files
## Experimental Design File
```{r}
# Import Environmental Data
design <- read.delim("../data/TC.design.txt", row.names = 1)
```

## Microbial Data
```{r}
# Import OTU data
# Import Raw Data
otu.in <- read.otu("../data/TC.bac.final.shared")

# Correct Sample IDs and Subset Original Design File
rownames(design) <- gsub("_0", "_", gsub("^ECU_", "", rownames(design)))
design.otu <- design[which(rownames(design) %in% rownames(otu.in)), ]
all.equal(rownames(otu.in), rownames(design.otu))

# Remove OTUs with less than two occurences across all sites
otus <- otu.in[, which(colSums(otu.in) >= 2)]

# Make Presence Absence Matrix
dataPA <- (otus > 0) * 1

# Import Taxonomy File
otu.tax <- read.tax(taxonomy = "../data/TC.bac.final.0.03.taxonomy",
                   format = "rdp", tax.levels = 6, col.tax = 3)
```

# Diversity Metrics - Hypothesis Testing - baseflow only (sed and water)
```{r Diversity Metrics - Bacteria, echo=TRUE}
# Rarefy Abundances (min abundance is 7172. We are sampling to 7100)
#baseflow only
df.design.otus <-cbind(design.otu,otus)
as.data.frame(df.design.otus)
df.baseflow <- subset(df.design.otus, Storm == "baseflow")

otus <- df.baseflow[,-c(1:6)]
design.baseflow <- df.baseflow[,c(1:6)]

min(rowSums(otus))
otus.r <- rrarefy(otus, 7100)

# Fisher's Alpha
fisher <- fisher.alpha(otus.r)

# Species Richness
#richness <- rowSums((PWESdata.r >= 1))
richness <- rowSums((otus.r >= 1))

# Shannon Diversity
shannon <- diversity(otus.r, "shannon")

# Simpson's Evenness
simp.even <- apply(otus.r, 1, simp_even)

#Pielou’s evenness
J <- shannon/log(specnumber(otus.r[,-c(1:1)]))

#combined richness, diversity, evenness
diversity <- cbind(design.baseflow,richness,shannon,simp.even,J)
#write.csv(diversity,"../data/diversity.bact.raw.csv")
```

#Plot OTU richness - baseflow - sed + water
```{r Plot - Shannon Diversity, echo=TRUE}
diversity2 <- diversity[diversity$Storm =="baseflow",]
baseflow.diversity <- diversity2[diversity2$Location !="A",]

# Graphing OTU richness
p <- ggplot(baseflow.diversity, aes(x=Location, y=richness, color=as.factor(Sample_Type))) + geom_boxplot() +
      geom_point(aes(color=factor(Sample_Type)), size=2, position = position_jitterdodge()) +   
      scale_color_manual(name="Sample Type", 
                         values=c("brown", "orange", "darkblue"), 
                         labels = c("stream bed", "stream bank", "water")) 
p + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "Location", y = "OTU richness") + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1)) +
    scale_x_discrete(breaks=c("B", "C", "D"), labels=c("downstream (B)", "middle (C)", "upstream (D)"))

ggsave("../figures/baseflow_OTUrichness_boxplots.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```

# Simple Hypothesis Testing Bacterial Community Composition - baseflow only
```{r}
#PERMANOVA on baseflow only

# Make Relative Abundence Matrix
otus <- df.baseflow[,-c(1:6)]
design.bf <- df.baseflow[,c(1:6)]

relabun <- otus
for (i in 1:dim(otus)[1]) {
  relabun[i, ] <- otus[i, ]/sum(otus[i,])
}

df.baseflow <-cbind(design.bf, relabun)

#df.baseflow
adonis = adonis(df.baseflow[,-c(1:6)] ~ Sample_Type, method = "bray", data = df.baseflow, perm=1000)
adonis
```
## Principal Coordinates Ordination - baseflow only
```{r}
# Principal Coordinates Analysis
sampleREL.dist <- vegdist(df.baseflow[,-c(1:6)], method="bray")

TC_pcoa <- cmdscale(sampleREL.dist, k=3, eig=TRUE, add=FALSE)
  # Classical (Metric) Multidimensional Scaling; returns PCoA coordinates
  # eig=TRUE returns eigenvalues; k = # of dimensions to calculate

explainvar1 <- round(TC_pcoa$eig[1] / sum(TC_pcoa$eig), 3) * 100
explainvar2 <- round(TC_pcoa$eig[2] / sum(TC_pcoa$eig), 3) * 100
sum.eig <- sum(explainvar1, explainvar2)
explainvar1 #29.5
explainvar2 #20.7

points <- cbind(as.data.frame(TC_pcoa$points), df.baseflow$Sample_Type)
pcoa.L.centroids <- melt(points, id="df.baseflow$Sample_Type", measure.vars = c("V1", "V2"))
pcoa.centroids <- cast(pcoa.L.centroids, variable ~ df.baseflow$Sample_Type, mean)
pcoa.centroids.se <- cast(pcoa.L.centroids, variable ~ df.baseflow$Sample_Type, se)
pcoa.centroids.sd <- cast(pcoa.L.centroids, variable ~ df.baseflow$Sample_Type, sd)
```

## Principal Coordinates Ordination - baseflow only
```{r Ordination (PCoA) - Bacteria, include=FALSE}
# Combine
trt <- c("BD","BK","W")
pcoa.cent.dataframe <- cbind(t(pcoa.centroids), t(pcoa.centroids.se))
colnames(pcoa.cent.dataframe) <- c("V1", "V2", "V1e", "V2e")
pcoa.cent.treats <- rownames(pcoa.cent.dataframe)
pcoa.cent.dataframe.trts <- as.data.frame(pcoa.cent.dataframe)
pcoa.cent.dataframe.trts$trt <- as.factor(trt)

#Plot
df1a <- as.data.frame(pcoa.cent.dataframe.trts)
plot1a <- ggplot(df1a, aes(x=V1, y=V2, colour=trt, shape = trt, label = trt)) + theme_bw() 
plot1a + theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank(), 
              axis.line = element_line(colour = "black")) + 
              theme(panel.background = element_blank()) + 
geom_point(aes(colour=trt), size=5, stroke = 1.25, show.legend = TRUE) + 
geom_text(hjust = 0, nudge_x = 0.05) +
geom_errorbarh(aes(xmax=V1+V1e, xmin=V1-V1e, height=0.01), colour="black") + 
geom_errorbar(aes(ymax=V2+V2e, ymin=V2-V2e, width=0.01), colour="black") +
scale_colour_manual(labels = c("stream bed", "stream bank", "water"), 
                    values = c("gray40", "gray40", "gray40")) + 
scale_shape_manual(labels = c("stream bed", "stream bank", "water"), values = c(17,15,19)) +
      theme(axis.title = element_text(size=14), axis.text = element_text(size=14),
          axis.text.x = element_text(size=14), 
          panel.border = element_rect(colour = "black", size = 1.25)) + 
      theme(axis.ticks.length = unit(0.3, "cm")) + 
          xlab("PCoA 1 (29.5%)") + ylab("PCoA 2 (20.7%)") +
          labs(shape = "Sample Type") +
          guides(shape = guide_legend(override.aes = list(size = 4, color="gray40")), color = "none")
ggsave("../figures/TC_ordination_type_baseflow.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```

# Bacterial community indicator species - baseflow only
```{r Bacteria Indicator Species, include=FALSE}
#df.baseflow <-cbind(design.bf, relabun)
design.type <- df.baseflow$Sample_Type
dataREL.t <- df.baseflow[,-c(1:6)]

dataREL <- dataREL.t[, colSums(dataREL.t) > 0.05]
bac.ind <- indval(dataREL, design.type)
levels(design.type)
summary(bac.ind)

inds <- which(bac.ind$pval <= 0.05)
bac.indicators <- as.data.frame(matrix(NA, nrow = length(inds), ncol = 4))
colnames(bac.indicators) <- c("OTU", "Cluster", "IndVal", "Prob")

bac.indicators$OTU <- names(inds)
bac.indicators$Cluster <- bac.ind$maxcls[inds]
bac.indicators$IndVal <- bac.ind$indcls[inds]
bac.indicators$Prob <- bac.ind$pval[inds]

ind.tax <- otu.tax[which(as.character(otu.tax$OTU) %in% bac.indicators$OTU), ]
ind.tax <- ind.tax[match(ind.tax$OTU, bac.indicators$OTU), ]

indicator.bac <- cbind(bac.indicators, ind.tax[, -c(1)])

indicator.bac <- indicator.bac[order(as.numeric(indicator.bac$Cluster)), ]

table(indicator.bac$Cluster)
table(indicator.bac$Phylum)
table(indicator.bac$Cluster)
levels(design.type)

# Export Bacteria Indicator Table
write.table(indicator.bac, "../data/BacterialIndicators_baseflow.txt",
            sep="\t", row.names = F, quote = F)
```

## Phylogenetic Tree and UniFrac Distance Matrix - did not run unifrac PCoA
```{r}
# Import Tree
tree <- read.tree("../data/TC.bac.rename.tree")

tree$tip.label[1:5]
tree$tip.label <- gsub("\\|", "", tree$tip.label)

# Import Unifrac Matrix
unifrac.raw <- read.delim("../data/TC.bac.tree1.weighted.phylip.dist",
                             header = F, skip = 1, row.names = 1)

row.names(unifrac.raw) <- gsub("     ", "", row.names(unifrac.raw))

unifrac <- unifrac.raw[which(row.names(unifrac.raw) %in%
                                   row.names(otus)),
                             which(row.names(unifrac.raw) %in%
                                   row.names(otus))]

rownames(unifrac) <- gsub("     ", "", row.names(unifrac))
colnames(unifrac) <- rownames(unifrac)
dim(unifrac)

# Make into Distance Matrix
unifrac.dist <- as.dist(unifrac, upper = T, diag = T)
```

## Notes: Phylogenetic Analysis
```{sh, eval = F}
# The following was done outside of R
python ../bin/name_change.py TC.bac.final.0.03.rep.fasta TC.bac.final.0.03.rep.rename.fasta

# This version is needed for Mothur
FastTree -gtr -nt -gamma -fastest TC.bac.final.0.03.rep.fasta > TC.bac.tree

# This version is needed for R analysis
FastTree -gtr -nt -gamma -fastest TC.bac.final.0.03.rep.rename.fasta > TC.bac.rename.tree

Mothur (v 1.39.5)
system(cp ./TC.trim.contigs.good.unique.good.filter.good.unique.precluster.pick.pick.opti_mcc.unique_list.0.03.rep.fasta ./TC.bac.final.0.03.rep.fasta)
system(cp ./TC.trim.contigs.good.unique.good.filter.good.unique.precluster.pick.pick.opti_mcc.unique_list.0.03.rep.count_table ./TC.bac.final.count)
unifrac.weighted(tree=TC.bac.tree, count=TC.bac.final.count, distance=square, processors=8)

Output File Names:
TC.bac.treewsummary
TC.bac.tree1.weighted.phylip.dist
```

# Phylogenetic Diversity
```{r Phylogenetic Diversity, include=FALSE}
# Reorder Tree
which(tree$edge.length < 0.0000001)
#tree.2 <- reorder(tree, order = "cladewise")
#tree.r <- midpoint.root(tree.2)
#phylo.d <- pd(otus.r, tree, include.root = F) #does this have to be rarefied? CHECK***

#otus <- df.baseflow[,-c(1:6)]
#design.baseflow <- df.baseflow[,c(1:6)]

phylo.d <- pd(otus, tree, include.root = F)

design.r <- design.baseflow[which(rownames(design.baseflow) %in% rownames(otus)), ]


design.bf.bcd <- design.r[design.r$Location != "A", ]

bf.bac <- as.data.frame(matrix(NA, dim(design.bf.bcd)[1], 10))
colnames(bf.bac) <- c("Sample", "Location", "Time", "Type", "Replicate", 
                       "S", "F_alpha", "S_even", "Shannon", "Phylo")
bf.bac$Sample <- rownames(design.bf.bcd)
bf.bac$Location <- design.bf.bcd$Location
bf.bac$Time <- design.bf.bcd$Storm
bf.bac$Type <- design.bf.bcd$Sample_Type
bf.bac$Replicate <- design.bf.bcd$Field_replicate
for(i in 1:length(rownames(design.bf.bcd))){
  temp <- rownames(design.bf.bcd)[i]
  bf.bac$S[i] <- richness[which(names(richness) == temp)]
  bf.bac$F_alpha[i] <- fisher[which(names(fisher) == temp)]
  bf.bac$S_even[i] <- simp.even[which(names(simp.even) == temp)]
  bf.bac$Shannon[i] <- shannon[which(names(shannon) == temp)]
  bf.bac$Phylo[i] <- phylo.d[which(rownames(phylo.d) == temp), 1]
}

# Reorganize
#bf.bac.m <- melt(bf.bac, id = c("Sample", "Location", "Time", "Type"))
#bf.bac.c <- cast(data = bf.bac.m, Location + Type ~ variable, c(mean, se), na.rm=T)

#bf.bac.c <- as.data.frame(bf.bac.c)
#bf.bac.m <- as.data.frame(bf.bac.m)

# # ANOVA - update
aov.sed=aov(Phylo~Location*Type,data=bf.bac)
summary(aov.sed)
posthocD <-TukeyHSD(aov.sed,"Location:Type",conf.level=0.95)
posthocD
```

# Phylo Diversity Plots
```{r}
#boxplots
#phylo.d <- pd(otus.r, tree, include.root = F)
#design.r <- design.bf[which(rownames(design.bf) %in% rownames(otus.r)), ]
#df.design.phylo <- cbind(design.r,phylo.d)

p <- ggplot(bf.bac, aes(x=Location, y=Phylo, color=as.factor(Type))) + geom_boxplot() +
      geom_point(aes(color=factor(Type)), size=2, position = position_jitterdodge()) +   
      scale_color_manual(name="Location", 
                         values=c("brown", "orange", "darkblue"), 
                         labels = c("stream bed", "stream bank", "water")) 
p + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "Location", y = "Phylogenetic Diversity") + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1)) +
    scale_x_discrete(breaks=c("B", "C", "D"), labels=c("downstream (B)", "middle (C)", "upstream (D)"))

ggsave("../figures/baseflow_PD_boxplots.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```

## Denitrification
```{r}
# Import Data
dat <- read.delim("../data/WQ_DEA_separate/2015_TC_DEArate.txt")
dat$Replicate <- as.factor(dat$Replicate)
```
# Denitrification Plots - baseflow only
```{r}
dat0 <- dat[dat$Time == "baseline", ]
dat1 <- dat0[dat0$Location != "A", ]
dat2 <- dat1[dat1$Location != "E", ]
dat.bf.bcd <- dat2[dat2$Location != "F", ]

dim1 <- length(dat.bf.bcd$acetyleneb[dat.bf.bcd$acetyleneb == "-"])

bf.eff <- as.data.frame(matrix(NA, dim1, 4))
colnames(bf.eff) <- c("Location", "Type", "Replicate", "Rate")
bf.eff$Location <- dat.bf.bcd$Location[dat.bf.bcd$acetyleneb == "-"]
bf.eff$Type <- dat.bf.bcd$Type[dat.bf.bcd$acetyleneb == "-"]
bf.eff$Replicate <- dat.bf.bcd$Replicate[dat.bf.bcd$acetyleneb == "-"]
bf.eff$Rate <- (dat.bf.bcd$Rate[dat.bf.bcd$acetyleneb == "+"] -
					   dat.bf.bcd$Rate[dat.bf.bcd$acetyleneb == "-"] )

# Reorganize
#bf.eff.m <- melt(bf.eff)
#bf.eff.c <- cast(data = bf.eff.m, Location + Type ~ variable, c(mean, se), na.rm=T)

#bf.eff.c <- as.data.frame(bf.eff.c)
#bf.eff.m <- as.data.frame(bf.eff.m)

# # ANOVA - update
aov.bf=aov(Rate~Location*Type,data=bf.eff)
summary(aov.bf)
posthocD <-TukeyHSD(aov.bf,"Location:Type",conf.level=0.95)
posthocD

p <- ggplot(bf.eff, aes(x=Location, y=Rate, color=as.factor(Type))) + geom_boxplot() +
      geom_point(aes(color=factor(Type)), size=2, position = position_jitterdodge()) +   
      scale_color_manual(name="Location", 
                         values=c("brown", "orange", "darkblue"), 
                         labels = c("stream bed", "stream bank", "water")) 
p + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "Location", 
          y = (expression(paste("Denitrification (ng N"[2],"O (g DM)"^-{1}," hr"^-{1},")")))) + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1)) +
    scale_x_discrete(breaks=c("B", "C", "D"), labels=c("downstream (B)", "middle (C)", "upstream (D)"))

ggsave("../figures/bf_denitrification_boxplot.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```
# Water Denitrification - baseflow + stormflow
```{r}
dat.water <- dat[dat$Type == "W", ]

# Deal with negative values
#dat.water[,7:10][dat.water[,7:10] < 0] <- 0

dim1 <- length(dat.water$acetyleneb[dat.water$acetyleneb == "-"])

wtr.eff <- as.data.frame(matrix(NA, dim1, 4))
colnames(wtr.eff) <- c("Location", "Time", "Replicate", "Rate")
wtr.eff$Location <- dat.water$Location[dat.water$acetyleneb == "-"]
wtr.eff$Time <- dat.water$Time[dat.water$acetyleneb == "-"]
wtr.eff$Replicate <- dat.water$Replicate[dat.water$acetyleneb == "-"]
wtr.eff$Rate <- (dat.water$Rate[dat.water$acetyleneb == "+"] -
					   dat.water$Rate[dat.water$acetyleneb == "-"] )

# Reorganize Data
#wtr.eff.m <- melt(wtr.eff)
#wtr.eff.c <- cast(data = wtr.eff.m, Location + Time ~ variable, c(mean, se), na.rm=T)

#wtr.eff.c <- as.data.frame(wtr.eff.c)
#wtr.eff.m <- as.data.frame(wtr.eff.m)

# # ANOVA - update
aov.wtr=aov(Rate~Location*Time,data=wtr.eff)
summary(aov.wtr)

p <- ggplot(wtr.eff, aes(x=Location, y=Rate, color=as.factor(Time))) + geom_boxplot() +
      geom_point(aes(color=factor(Time)), size=2, position = position_jitterdodge()) +   
      scale_color_manual(name="Location", 
                         values=c("cyan", "darkblue"), 
                         labels = c("baseflow", "stormflow")) 
p + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "Location", 
          y = (expression(paste("Denitrification (ng N"[2],"O (g DM)"^-{1}," hr"^-{1},")")))) + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1))
    #scale_x_discrete(breaks=c("A","B","C", "D","E","F"), labels=c("downstream (B)", "middle (C)", "upstream (D)"))

ggsave("../figures/water_denitrification_boxplot.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```

# Diversity Function Plots
```{r}
#graph baseflow all sample types only - graph PD vs denitrification, graph OTU richness vs denitrification
bf.phylo.fxn.full <- cbind(bf.eff, bf.bac)
bf.phylo.fxn <- bf.phylo.fxn.full[,-c(1:3)]

#graph denitrification x alpha diversity (PD)
p <- ggplot(bf.phylo.fxn, aes(x=Phylo, y=Rate, shape=as.factor(Location), color=as.factor(Type))) +
    scale_color_manual(name="Sample Type", 
                      labels = c("stream bed", "stream bank", "water"), 
                      values = c("brown","orange", "darkblue")) +
                      geom_point(size=4) +
    scale_shape_manual(name="Location", 
                       labels = c("downstream (B)", "middle (C)", "upstream (D)"), 
                       values = c(17,15,19))
denitrif.phylo <- p + 
              theme_bw() + 
              theme(panel.grid.major = element_blank(), 
                    panel.grid.minor = element_blank(), 
                    axis.line = element_line(colour = "black")) +
              theme(axis.title=element_text(vjust=1,size=16), 
                    axis.text=element_text(size=16), 
                    axis.text.x = element_text(vjust=0.65, hjust=0.5, size=16), 
                    panel.border = element_rect(colour = "black",size=1)) + 
              theme(axis.ticks.length=unit(0.3,"cm")) + 
              labs(x = "Phylogenetic Diversity", 
                   y = (expression(paste("Denitrification (ng N"[2],"O (g DM)"^-{1}," hr"^-{1},")")))) +   
              theme(strip.text.x = element_text(size=16, face="bold"), 
                    strip.text.y = element_text(size=16, face="bold"), 
                    strip.background = element_rect(colour="black", fill="white", size=1))
denitrif.phylo
ggsave("../figures/baseflow_denitrif_PD.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)

#graph denitrification x alpha diversity (OTU richness)
p <- ggplot(bf.phylo.fxn, aes(x=S, y=Rate, shape=as.factor(Location), color=as.factor(Type))) +
    scale_color_manual(name="Sample Type", 
                      labels = c("stream bed", "stream bank", "water"), 
                      values = c("brown","orange", "darkblue")) +
                      geom_point(size=4) +
    scale_shape_manual(name="Location", 
                       labels = c("downstream (B)", "middle (C)", "upstream (D)"), 
                       values = c(17,15,19))
denitrif.OTUrichness <- p + 
              theme_bw() + 
              theme(panel.grid.major = element_blank(), 
                    panel.grid.minor = element_blank(), 
                    axis.line = element_line(colour = "black")) +
              theme(axis.title=element_text(vjust=1,size=16), 
                    axis.text=element_text(size=16), 
                    axis.text.x = element_text(vjust=0.65, hjust=0.5, size=16), 
                    panel.border = element_rect(colour = "black",size=1)) + 
              theme(axis.ticks.length=unit(0.3,"cm")) + 
              labs(x = "OTU Richness", 
                   y = (expression(paste("Denitrification (ng N"[2],"O (g DM)"^-{1}," hr"^-{1},")")))) +   
              theme(strip.text.x = element_text(size=16, face="bold"), 
                    strip.text.y = element_text(size=16, face="bold"), 
                    strip.background = element_rect(colour="black", fill="white", size=1))
denitrif.OTUrichness
ggsave("../figures/baseflow_denitrif_OTUrichness.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)

#graph denitrification x alpha diversity (PD)
p <- ggplot(bf.phylo.fxn, aes(x=Phylo, y=S, shape=as.factor(Location), color=as.factor(Type))) +
    scale_color_manual(name="Sample Type", 
                      labels = c("stream bed", "stream bank", "water"), 
                      values = c("brown","orange", "darkblue")) +
                      geom_point(size=4) +
    scale_shape_manual(name="Location", 
                       labels = c("downstream (B)", "middle (C)", "upstream (D)"), 
                       values = c(17,15,19))
phylo.OTUrichness <- p + 
              theme_bw() + 
              theme(panel.grid.major = element_blank(), 
                    panel.grid.minor = element_blank(), 
                    axis.line = element_line(colour = "black")) +
              theme(axis.title=element_text(vjust=1,size=16), 
                    axis.text=element_text(size=16), 
                    axis.text.x = element_text(vjust=0.65, hjust=0.5, size=16), 
                    panel.border = element_rect(colour = "black",size=1)) + 
              theme(axis.ticks.length=unit(0.3,"cm")) + 
              labs(x = "Phylogenetic Diversity", y = "OTU Richness") +   
              theme(strip.text.x = element_text(size=16, face="bold"), 
                    strip.text.y = element_text(size=16, face="bold"), 
                    strip.background = element_rect(colour="black", fill="white", size=1))
phylo.OTUrichness
ggsave("../figures/bf_phylo_OTUrichness.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```

# Diversity Metrics - Hypothesis Testing - baseflow only (sed and water)
```{r Diversity Metrics - Bacteria, echo=TRUE}
# Rarefy Abundances (min abundance is 7172. We are sampling to 7100)
#baseflow and stormflow
#df.design.otus <-cbind(design.otu,otus)
#as.data.frame(df.design.otus)
df.basestorm <- subset(df.design.otus, Sample_Type == "W")
str(df.basestorm)

otus <- df.basestorm[,-c(1:6)]
design.basestorm <- df.basestorm[,c(1:6)]

rowSums(otus)
#TC_48 68; TC_49 405; TC_50 226; TC_51 328

# OTU table - removed low abundance samples
otus.w <- otus[which(rowSums(otus) >= 100), ]
dim(otus.w)

min(rowSums(otus.w))
otus.w.r <- rrarefy(otus.w, 200)

# Fisher's Alpha
fisher.w <- fisher.alpha(otus.w.r)

# Species Richness
richness.w <- rowSums((otus.w >= 1))

# Shannon Diversity
shannon.w <- diversity(otus.w.r, "shannon")

# Simpson's Evenness
simp.even.w <- apply(otus.w.r, 1, simp_even)

#Pielou’s evenness
J.w <- shannon/log(specnumber(otus.w.r[,-c(1:1)]))

# OTU table - odd sites in bacterial composition data and remove in design file
odd.sites.w <- c("TC_48")

otu_final <- otus.w[setdiff(rownames(otus.w), odd.sites.w), ]
design_final <- design.basestorm[setdiff(rownames(design.basestorm), odd.sites.w), ]
bfsf.basestorm <- cbind(design_final, otu_final)

#combined richness, diversity, evenness
basestorm.diversity <- cbind(design_final,richness.w,shannon.w,simp.even.w,J.w)
```

#Plot OTU richness - basestorm - water
```{r Plot - OTU richness, echo=TRUE}
# Graphing OTU richness
p <- ggplot(basestorm.diversity, aes(x=Location, y=richness.w, color=as.factor(Storm))) + geom_boxplot() +
      geom_point(aes(color=factor(Storm)), size=2, position = position_jitterdodge()) +   
      scale_color_manual(name="Flow", 
                         values=c("cyan", "darkblue"), 
                         labels = c("baseflow", "stormflow")) 
p + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "Location", y = "OTU richness") + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1)) 
#+scale_x_discrete(breaks=c("B", "C", "D"), labels=c("downstream (B)", "middle (C)", "upstream (D)"))

ggsave("../figures/basestorm_OTUrichness_boxplots.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```

# Simple Hypothesis Testing Bacterial Community Composition - baseflow + stormflow only
```{r}
#PERMANOVA on base+stormflow
#bfsf.basestorm <- cbind(design_final, otu_final)

# Make Relative Abundence Matrix
otus <- bfsf.basestorm[,-c(1:6)]
design.bfsf <- bfsf.basestorm[,c(1:6)]

# Remove OTUs with less than two occurences across all sites
otus.bfsf <- otus[, which(colSums(otus) >= 2)]

relabun.w <- otus.bfsf
for (i in 1:dim(otus.bfsf)[1]) {
  relabun.w[i, ] <- otus.bfsf[i, ]/sum(otus.bfsf[i,])
}

df.basestorm <-cbind(design.bfsf, relabun.w)

#df.basestorm
adonis = adonis(df.basestorm[,-c(1:6)] ~ Storm * Location, method = "bray", data = df.basestorm, perm=1000)
adonis
```
## Principal Coordinates Ordination - base+stormflow water
```{r}
# Principal Coordinates Analysis
sampleREL.dist <- vegdist(df.basestorm[,-c(1:6)], method="bray")

TC_pcoa <- cmdscale(sampleREL.dist, k=3, eig=TRUE, add=FALSE)
  # Classical (Metric) Multidimensional Scaling; returns PCoA coordinates
  # eig=TRUE returns eigenvalues; k = # of dimensions to calculate

explainvar1 <- round(TC_pcoa$eig[1] / sum(TC_pcoa$eig), 3) * 100
explainvar2 <- round(TC_pcoa$eig[2] / sum(TC_pcoa$eig), 3) * 100
sum.eig <- sum(explainvar1, explainvar2)
explainvar1 #28.7
explainvar2 #22.4

pcoa.groups <- paste(design.bfsf$Storm, design.bfsf$Location, sep = "_")

pcoa.points <- data.frame(TC_pcoa$points, group = pcoa.groups)

# Calculate Centroids (mean and SE)
pcoa.L.centroids <- melt(pcoa.points, id="group", measure.vars = c("X1", "X2"))
pcoa.centroids <- cast(pcoa.L.centroids, variable ~ group, mean)
pcoa.centroids.se <- cast(pcoa.L.centroids, variable ~ group, se)
pcoa.centroids.sd <- cast(pcoa.L.centroids, variable ~ group, sd)
```

## Principal Coordinates Ordination - base+stormflow - water
```{r Ordination (PCoA) - Bacteria, include=FALSE}
# Combine
pcoa.cent.dataframe <- cbind(t(pcoa.centroids), t(pcoa.centroids.se))
colnames(pcoa.cent.dataframe) <- c("V1", "V2", "V1e", "V2e")
pcoa.cent.treats <- rownames(pcoa.cent.dataframe)

trt.flow <- c("baseflow","baseflow","baseflow","baseflow",
              "stormflow","stormflow","stormflow","stormflow","stormflow","stormflow","stormflow","stormflow")
trt.loc <- c("A","B","C","D","A","B","C","D","E","F","G","I")
pcoa.cent.dataframe.trts <- as.data.frame(pcoa.cent.dataframe)
pcoa.cent.dataframe.trts$trt.flow <- as.factor(trt.flow)
pcoa.cent.dataframe.trts$trt.loc <- as.factor(trt.loc)

#Plot
df1a <- as.data.frame(pcoa.cent.dataframe.trts)
plot1a <- ggplot(df1a, aes(x=V1, y=V2, colour=df1a$trt.flow, shape = df1a$trt.loc, label=trt.loc)) + theme_bw() 
plot1a + theme(panel.grid.major = element_blank(), 
              panel.grid.minor = element_blank(), 
              axis.line = element_line(colour = "black")) + 
              theme(panel.background = element_blank()) + 
geom_point(aes(colour=trt.flow), size=5, stroke = 1.25, show.legend = TRUE) + geom_text(hjust = 0, nudge_x = 0.05) +
geom_errorbarh(aes(xmax=V1+V1e, xmin=V1-V1e, height=0.01), colour="black") + 
geom_errorbar(aes(ymax=V2+V2e, ymin=V2-V2e, width=0.01), colour="black") +
scale_colour_manual(labels = c("baseflow", "stormflow"), 
                    values = c("cyan", "darkblue")) + 
scale_shape_manual(labels = c("A","B","C","D","E","F","G","I"), 
                   values = c(15,16,17,18,24,21,22,23)) +
      theme(axis.title = element_text(size=14), axis.text = element_text(size=14),
          axis.text.x = element_text(size=14), 
          panel.border = element_rect(colour = "black", size = 1.25)) + 
      theme(axis.ticks.length = unit(0.3, "cm")) + 
          xlab("PCoA 1 (28.7%)") + ylab("PCoA 2 (22.4%)") +
          labs(shape = "Location", colour = "Storm")
          #guides(shape = guide_legend(override.aes = list(size = 4, color="gray70")), color = "none")
ggsave("../figures/TC_ordination_water_basestorm.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```

# Phylogenetic Diversity - water only
```{r Phylogenetic Diversity, include=FALSE}
# Reorder Tree
which(tree$edge.length < 0.0000001)
#phylo.d <- pd(otus.r, tree, include.root = F) ##CHECK - use rarefied?
#df.basestorm <-cbind(design.bfsf, relabun.w) #water only

phylo.w.d <- pd(otus.bfsf, tree, include.root = F)

design.w.r <- design.bfsf[which(rownames(design.bfsf) %in% rownames(otus.bfsf)), ]

bfsf.bac <- as.data.frame(matrix(NA, dim(design.w.r)[1], 10))
colnames(bfsf.bac) <- c("Sample", "Location", "Time", "Type", "Replicate", 
                       "S", "F_alpha", "S_even", "Shannon", "Phylo")
bfsf.bac$Sample <- rownames(design.w.r)
bfsf.bac$Location <- design.w.r$Location
bfsf.bac$Time <- design.w.r$Storm
bfsf.bac$Type <- design.w.r$Sample_Type
bfsf.bac$Replicate <- design.w.r$Field_replicate
for(i in 1:length(rownames(design.w.r))){
  temp <- rownames(design.w.r)[i]
  bfsf.bac$S[i] <- richness[which(names(richness.w) == temp)]
  bfsf.bac$F_alpha[i] <- fisher[which(names(fisher.w) == temp)]
  bfsf.bac$S_even[i] <- simp.even[which(names(simp.even.w) == temp)]
  bfsf.bac$Shannon[i] <- shannon[which(names(shannon.w) == temp)]
  bfsf.bac$Phylo[i] <- phylo.d[which(rownames(phylo.w.d) == temp), 1]
}

# Reorganize
#bfsf.bac.m <- melt(bfsf.bac, id = c("Sample", "Location", "Time", "Type"))
#bfsf.bac.c <- cast(data = bfsf.bac.m, Location + Type ~ variable, c(mean, se), na.rm=T)

#bfsf.bac.c <- as.data.frame(bfsf.bac.c)
#bfsf.bac.m <- as.data.frame(bfsf.bac.m)

# # ANOVA - update
aov.wtr=aov(S~Location,data=bfsf.bac)
summary(aov.wtr)
posthocD <-TukeyHSD(aov.wtr,"Location",conf.level=0.95)
posthocD
```

# Phylo Diversity Plots
```{r}
#boxplots
p <- ggplot(bfsf.bac, aes(x=Location, y=Phylo, color=as.factor(Time))) + geom_boxplot() +
      geom_point(aes(color=factor(Time)), size=2, position = position_jitterdodge()) +   
      scale_color_manual(name="Time", 
                         values=c("cyan", "darkblue"), 
                         labels = c("baseflow", "stormflow")) 
p + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "Location", y = "Phylogenetic Diversity") + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1)) 

ggsave("../figures/basestorm_PD_boxplots.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```